---
title: "netsum-overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{netsum-overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

`netsum` is an R package and Shiny web application designed to make network analysis more accessible for psychological and behavioral data.\
It provides a point-and-click interface for estimating, visualizing, and interpreting networks without needing to write R code.

This vignette serves as a quick overview and usage guide.

------------------------------------------------------------------------

## Features

-   Upload CSV files with automatic validation (checks for format and missing values)
-   Select variables to include as nodes in the network
-   Choose from multiple network estimation methods (`EBICglasso`, `glasso`, `pcor`, `cor`)
-   Group-wise network estimation (optional)
-   Fully customizable network plots
    -   Custom plot title, layout, and label color
    -   Dynamic node grouping and color selection
-   Live network summary messages (e.g., success warnings, empty edges)
-   Launch the app with a single function: `netsum::launch_netsum_app()`

------------------------------------------------------------------------

## Installation and example work flow

To install `netsum` from a local folder:

```{r, eval = FALSE}
install.packages(".", repos = NULL, type = "source")
```

After installation, launch the app using:

```{r, eval = FALSE}
library(netsum)
launch_netsum_app()
```

This will open the interactive app in your default web browser.

## Using the App

Once launched:

1.  Upload a .csv file with numeric variables

2.  Click "Check Data" to verify format and missing values

3.  Select variables to include as network nodes

4.  Optionally specify a grouping variable (e.g., by condition)

5.  Choose a network estimation method

6.  Set a tuning parameter if applicable (only for EBICglasso and glasso")

7.  Click "Estimate Network"


You will see the network plot update in real time. You can also:

-   Add group names and node ranges (e.g., 1:3;4:6)

-   Pick custom colors dynamically for each group

-   Adjust label color, layout, and title


## Notes

The app is built using `shiny`, `bootnet`, `qgraph`, and `colourpicker`.

Network plots are drawn using the `plot_network()` function.

Estimation logic is handled by `compute_network()`.

The app dynamically adapts based on user input (e.g., color pickers appear only if groups are named).



## Authors and Credits

This vignette is part of the the *Programming the Next Step 2025* course.  
Developed by Lucas Straub.

